Forward Sensitivity Equations in the Presence of Events

### 2018-11-30 paper.tex
Try to make the text more clear

\paragraph{Replacement} 
The value of the $i$'th state, $x_i$, is set to a predefined value $v_{e, i}$. In that case, the derivative vanishes and the sensitivity at time $t_e$ is
\begin{align}
    \lim_{t\searrow t_e}\frac{\partial x_i}{\partial t_e} = \left.\frac{\partial f_i}{\partial x}\right|_{x_e}\cdot (x_e - v_e) - f_i(x_e),
    \label{eq:replace}
\end{align}
where we have used the definition of $b(t_e)$. In the equation, $\cdot: \mathbb R^n\times \mathbb R^n\rightarrow \mathbb R$ denotes the scalar product. While the state value $x_i$ jumps, the other states $x_j$, $j\neq i$, are continued. Continuation means that $v_{e, j}$ is set to $x_{e, j}$ and we need eq.~\eqref{eq:leqte} evaluated at $t_e$ to get $\frac{\partial v_e}{\partial t_e} = Ax_e + b(t_e)$. The sensitivities become 
\begin{align}
    \lim_{t\searrow t_e}\frac{\partial x_j}{\partial t_e} = \left.\frac{\partial f_j}{\partial x}\right|_{x_e}\cdot (x_e - v_e).\label{eq:continued}
\end{align}
The contributions from $b$ cancel out. In conclusion, we find that an event at time $t_e$ affects sensitivities with respect to $t_e$ of both the affected and unaffected states.\\

\paragraph{Additive} 
A constant $\Delta x_i$ is added to the value of the $i$'th state, $x_i$. In that case, $v_e$ is set to eq.~\eqref{eq:leqte} evaluated at $t_e$ plus the constant $\Delta x_i$. Therefore, the same argumentation as in the continued case holds and the sensitivity is the same as eq.~\eqref{eq:continued}. The difference $x_e - v_e$ turns out to be either 0 (for the continued states) or $-\Delta x_i$ (for the affected state $i$).\\

\paragraph{Multiplicative}
The value of the $i$'th state, $x_i$, is multiplied with a constant $\alpha_i$. Eq.~\eqref{eq:leqte} is evaluated at $t_e$ and multiplied by $\alpha$ to get $v_e$. Consequently, the derivative $\frac{\partial v_e}{\partial t_e}$ is computed and plugged into eq.~\eqref{eq:jump}, yielding
\begin{align}
    \lim_{t\searrow t_e}\frac{\partial x_i}{\partial t_e} = \left.\frac{\partial f_i}{\partial x}\right|_{x_e}\cdot (x_e - v_e) - (1-\alpha_i)f_i(x_e).
\end{align}
The difference $x_e - v_e$ is either 0 (for the continued states) or $(1-\alpha_i)x_{e, i}$ (for the multiplied state).

@@

Without loss of generality, let ``var'' be $x_1$. The time and value parameters are denoted as $\tau$ and $\xi$, and the value be replaced. This event with the required additional events is shown in Table~\ref{tab:replacement}.
\begin{table}[ht]
\caption{Replacement event and list of required additional sensitivity events.}
\label{tab:replacement}
\begin{center}
\begin{tabular}{|r|l|l|l|l|}\hline
     1 & $x_1$ & $\tau$ & $\xi$ & replace \\
     \hline
     2 & $\sfrac{\partial x_1}{\partial p_j}$ & $\tau$ & $0$ & replace \\
     3 & $\sfrac{\partial x_1}{\partial \tau}$ & $\tau$ & $J_{11}(x_1-\xi)-f_1$ & add \\
     4 & $\sfrac{\partial x_k}{\partial \tau}$ & $\tau$ & $J_{k1}(x_1-\xi)$ & add \\
     5 &$\sfrac{\partial x_1}{\partial \xi}$ & $\tau$ & 1 & add \\
     \hline
\end{tabular}
\end{center}
\end{table}
In the table, $j$ ranges from 1 to $m$, $J_{11}$ denotes the (1, 1)-element of the Jacobian $J_{ij} = \lim\limits_{t \nearrow \tau}\frac{\partial f_i}{\partial x_j}(x(t), p, t)$ and $f_1$ denotes the first element of $f_i := \lim\limits_{t \nearrow \tau}f_i(x(t), p, t)$. Note that, depending on the numeric implementation of the events, the symbolic expressions for $J_{11}$ and $f_1$ can be used and evaluated within the event function based on the state values at the event time before execution of the event.

@@

>>>>

\paragraph{Replacement}
The value of the $i$'th state, $x_i$, is set to a predefined value $v_{e, i}$. In that case, the derivative vanishes and the sensitivity at time $t_e$ is
\begin{align}
    \lim_{t\searrow t_e}\frac{\partial x_i}{\partial t_e} = \left.\frac{\partial f_i}{\partial x}\right|_{x_e}\cdot (x_e - v_e) - f_i(x_e),
    \label{eq:replace}
\end{align}
where we have used the definition of $b(t_e)$. In the equation, $\cdot: \mathbb R^n\times \mathbb R^n\rightarrow \mathbb R$ denotes the scalar product. While the state value $x_i$ jumps, the other states $x_j$, $j\neq i$, are continued. Continuation means that $v_{e, j}$ is set to $x_{e, j}$ and we need eq.~\eqref{eq:leqte} evaluated at $t_e$ to get $\frac{\partial v_e}{\partial t_e} = Ax_e + b(t_e)$. The sensitivities become
\begin{align}
    \lim_{t\searrow t_e}\frac{\partial x_j}{\partial t_e} = \left.\frac{\partial f_j}{\partial x}\right|_{x_e}\cdot (x_e - v_e).\label{eq:continued}
\end{align}
The contributions from $b$ cancel out. Due to the fact that all states except for $x_i$ are continued, the scalar product reduces to a single term. However, we find that an event at time $t_e$ affects sensitivities with respect to $t_e$ of both the affected and unaffected states.\\

\paragraph{Additive}
A constant $\Delta x_i$ is added to the value of the $i$'th state, $x_i$. In that case, $v_e$ is set to eq.~\eqref{eq:leqte} evaluated at $t_e$ plus the constant $\Delta x$. Therefore, the same argumentation as in the continued case holds for both, the affected and unaffected states, and the sensitivities are as eq.~\eqref{eq:continued} for states $i$ and $j\neq i$. \\

\paragraph{Multiplicative}
The value of the $i$'th state, $x_i$, is multiplied with a constant $\alpha_i$. Eq.~\eqref{eq:leqte} is evaluated at $t_e$ and multiplied by $\alpha$ to get $v_e$. Consequently, the derivative $\frac{\partial v_e}{\partial t_e}$ is computed and plugged into eq.~\eqref{eq:jump}, yielding
\begin{align}
    \lim_{t\searrow t_e}\frac{\partial x_i}{\partial t_e} = \left.\frac{\partial f_i}{\partial x}\right|_{x_e}\cdot (x_e - v_e) - (1-\alpha_i)f_i(x_e).
\end{align}
Same as for replacement events, the sensitivity of the affected and the unaffected states need to be distinguished. The term $(1-\alpha_i)f_i(x_e)$ only occurs for the affected state. Futhermore, all states except $x_i$ are continued. The scalar product therefore reduces to the $i$'th contribution.

@@

Without loss of generality, let ``var'' be $x_1$. The time and value parameters are denoted as $\tau$ and $\xi$, and the value be replaced. This event with the required additional events is shown in Table~\ref{tab:replacement}.
\begin{table}[ht]
\caption{Replacement event and list of required additional sensitivity events.}
\label{tab:replacement}
\begin{center}
\begin{tabular}{|r|l|l|l|l|}\hline
     1 & $x_1$ & $\tau$ & $\xi$ & replace \\
     \hline
     2 & $\sfrac{\partial x_1}{\partial p_j}$ & $\tau$ & $0$ & replace \\
     3 & $\sfrac{\partial x_1}{\partial \tau}$ & $\tau$ & $J_{11}(x_1-\xi)-f_1$ & add \\
     4 & $\sfrac{\partial x_k}{\partial \tau}$ & $\tau$ & $J_{k1}(x_1-\xi)$ & add \\
     5 &$\sfrac{\partial x_1}{\partial \xi}$ & $\tau$ & 1 & add \\
     \hline
\end{tabular}
\end{center}
\end{table}
In the table, $j$ ranges from 1 to $m$, $J_{11}$ denotes the (1, 1)-element of the Jacobian $J_{ij} = \lim\limits_{t \nearrow \tau}\frac{\partial f_i}{\partial x_j}(x(t), p, t)$ and $f_1$ denotes the first element of $f_i := \lim\limits_{t \nearrow \tau}f_i(x(t), p, t)$.
The differences $x_e - v_e$ are all zero except for the first state. Therefore, the scalar products in eqs.~\eqref{eq:replace} and \eqref{eq:continued} reduce to one contribution.
Note that, depending on the code implementation of the events, the symbolic expressions for $J_{11}$ and $f_1$ can be used and evaluated within the event function based on the state values at the event time \textbf{before} execution of the event.

@@

% Here we could discuss that different events have different \tau and \xi parameters that are reduced to one by a parameter estimation. Then the sensitivities would just be added wich corresponds to adding the jumps.
